\documentclass[12pt]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage{tikz}
\usepackage{tabto}
\usetikzlibrary{automata,positioning}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usetikzlibrary{automata,positioning}

\begin{document}

\title{CS4384 : Automata Theory\\Homework Assignment 2}
\author{Matthew McMillian\\mgm160130@utdallas.edu}
\maketitle


\begin{enumerate}
	
	\item Let L = $\{$ $0^i1^j0^k$ $|$ $k > i + j$ $\}$. Use the pumping lemma to show that L is not regular. \\ \\
	Assume L is regular and let p be it's pumping length. Define $S$ such that $S$ = $0^p1^p0^{2p+1}$, such that $|$S$| > p$ and that $S$ $\in$ L. Let $S$ = $xyz$ be a partitioning satisfying $|xy| \leq$ p, and $|y|$ $>$ 0. Since we know $|y| >$ 0, we know that $y$ contains only zeros. Since $y$ contains some number of zeros equal to $|y|$ = k for some k $\in$ (0,p], we can conclude that $xy$ contains $0^{p-k}$ zeros. Let $i$ = 2 such that $S^{'}$ = $xy^iz$ = $xy^2z$ = $xyyz$ = $0^{p-k}0^k0^k1^p0^{2p+1}$. Simplifying we obtain $0^{p+k}1^p0^{2p+1}$. This contradicts the pumping lemma, since we have come up with a string $S^{'}$ with pumping length p that is NOT in the language L since $k < i+j$ for $S^{'}$. Thus, this language is not regular. \\ \\
	
	\item Let L = $\{$ $w$ $|$ $w \in \{0,1\}^*$ $w$ is not a palindrome $\}$. Use the pumping lemma to show that L is not regular. \\ \\
	By closure properties of regular languages, if L is regular then $\overline{L}$ is also regular. Take for consideration the compliment of L, such that $\overline{L}$ = $\{$ $w$ $|$ $w \in \{0,1\}^*$ $w$ is a palindrome $\}$. Assume $\overline{L}$ is regular and let p be it's pumping length. Define $S$ such that $S$ = $0^p1^p0^p$, such that $|$S$| > p$ and that $S$ $\in$ $\overline{L}$. Let $S$ = $xyz$ be a partitioning satisfying $|xy| \leq$ p, and $|y|$ $>$ 0. Since we know $|y| >$ 0, we know that $y$ contains only zeros. Since $y $contains some number of zeros equal to $|y|$ = k for some k $\in$ (0,p], we can conclude that $xy$ contains $0^{p-k}$ zeros. Let $i$ = 2 such that $S^{'}$ = $xy^iz$ = $xy^2z$ = $xyyz$ = $0^{p-k}0^k0^k1^p0^{p}$. Simplifying we obtain $0^{p+k}1^p0^p$. This contradicts the pumping lemma, since we have come up with a string $S^{'}$ with pumping length p that is NOT in the language $\overline{L}$ since $S^{'}$ is not a palindrome because k is strictly greater than zero. Thus since $\overline{L}$ is not regular, the language L is not regular.
	\pagebreak
	
	\item Let L = $\{$ $0^m1^n$ $|$ $m \neq n$ $\}$. Use closure properties of regular languages to prove that L is not regular. \\ \\
	By closure properties of regular languages, if L is regular then $\overline{L}$ is also regular. Take for consideration the compliment of L, such that $\overline{L}$ = $\{$ $0^m1^n$ $|$ $m = n$ $\}$. Define $S$ such that $S$ = $0^p1^p$, such that $|$S$| > p$ and that $S$ $\in$ $\overline{L}$. Assume $\overline{L}$ is regular and let p be its pumping length. Let $S$ = $xyz$ be a partitioning satisfying $|xy| \leq$ p, and $|y|$ $>$ 0. Since we know $|y| >$ 0, we know that $y$ contains only zeros. Since $y $ contains some number of zeros equal to $|y|$ = k for some k $\in$ (0,p], we can conclude that $xy$ contains $0^{p-k}$ zeros.  Let $i$ = 2 such that $S^{'}$ = $xy^iz$ = $xy^2z$ = $xyyz$ = $0^{p-k}0^k0^k1^p$. Simplifying we obtain $0^{p+k}1^p$. This contradicts the pumping lemma, since we have come up with a string $S^{'}$ with pumping length p that is NOT in the language $\overline{L}$ since $m \neq n$ since k is greater than zero. Thus since $\overline{L}$ is not regular, the language L is not regular. \\ \\

	\item Let $\Sigma$ = $\{$ $0,1,+,=$ $\}$. Let ADD = $\{$ $x = y + z$ $|$ $x,y,z$ are binary integers, and x is the sum of y and z $\}$. Use the pumping lemma to prove ADD is not regular. \\ \\
	Assume ADD is regular and let p be its pumping length. Define $S$ such that $S$ = $1^{p+1} = 10^p + 1^p$, such that $|$S$| > p$ and that $S$ $\in$ L. Let $S$ = $xyz$ be a partitioning satisfying $|xy| \leq$ p, and $|y|$ $>$ 0. Since we know $|y| >$ 0, we know that $y$ contains only ones. Since $y$ contains some number of ones equal to $|y|$ = k for some k $\in$ (0,p], we can conclude that $xy$ contains $1^{p-k}$ ones. Let $i$ = 0 such that $S^{'}$ = $xy^iz$ = $xy^0z$ = $xz$ = $1^{p+1-k} = 10^p + 1^p$. This contradicts the pumping lemma, since we have come up with a string $S^{'}$ with pumping length p that is NOT in the language L since k is greater than 0; $S{'}$ will never be true. Thus, we can conclude that ADD is not regular. \\ \\
	
	\item Let L = $\{$ $1^ky$ $|$ $y \in \{0,1\}^*$ and y contains at least k 1s for k $\geq$ 1 $\}$. Use the pumping lemma to prove L is not regular. \\ \\ 
	Assume L is regular and let p be its pumping length.  Define $S$ such that $S$ = $1^p1^p$, such that $|$S$| > p$ and that $S$ $\in$ L. Let $S$ = $xyz$ be a partitioning satisfying $|xy| \leq$ p, and $|y|$ $>$ 0. Since we know $|y| >$ 0, we know that $y$ contains only ones. Since $y$ contains some number of ones equal to $|y|$ = k for some k $\in$ (0,p], we can conclude that $xy$ contains $1^{p-k}$ ones. Let $i$ = 2 such that $S^{'}$ = $xy^iz$ = $xy^2z$ = $xyyz$ = $1^{p-k}1^k1^k1^p$. Simplifying we obtain $1^{p+k}1^p$. This contradicts the pumping lemma, since we have come up with a string $S^{'}$ with pumping length p that is NOT in the language L since y does not have at least 'k' ones. Thus, this language is not regular.

\end{enumerate}

\end{document}
